Finding a closest point in a lattice of Voronoi's first kind

### 2014-07-08 reviews/response.tex

\section*{Reviewer 2}\label{sec:reviewer-2}

\begin{enumerate}
\item\textbf{Comment:}  
page 1, last line "The closest lattice point problem is known to be NP-hard under certain condition ...": I found this paragraph very confusing, and somehow inaccurate. What "certain conditions"? I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following. CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance. Interestingly (see 35 and 17) there are families of lattices when CVP remain NP-hard even when given the best possible basis (or any hard-to-compute side information about the lattice). This version of the problem is called CVP with preprocessing (CVPP), and it is particularly relevant to this work, where it is shown that CVPP can be solved exactly and in polynomial time for the class of lattices of Voronoi first kind, when given as auxiliary information a special basis (namely, an obtuse superbasis.) 
\\\textbf{Response:}

\item\textbf{Comment:}  
Additional comments about the above paragraph: one author is missing in the reference for 12. Of [1,23,40] only 23 achieves $n^{O(n)}$ running time. (I think the algorithms in 40 take $2^{O(n^2)}$.) There is a journal version of 37 in SIAM J. Computing, and the paper is by two authors (not just Micciancio). 
\\\textbf{Response:}

\item\textbf{Comment:}  
page 2, second paragraph: Aren't the results of 3,29 and 30 subsumed by 34? (The authors sure know better than me. A a short clarification may direct the interested reader to the best paper to read.) Also, CVP is always solvable in polynomial time when the dimension is fixed. So, it is unclear the significance of CVP algorithms for Leech lattice (in fixed dimension 24) in the context of this discussion. 
\\\textbf{Response:}
This sentence has been reworded and the reference list has been reduced.  It now reads:

``Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.''

The interested reader could read these references in the order given. %, i.e., first Chapter 4 of~\cite{SPLAG} and~\cite{Conway1982FastQuantDec} for $D_n$ and $D_n^*$ and older algorithms for $A_n$ and $A_n^*$.  Then~\cite{McKilliam2008} for a fast algorithm for $A_n^*$, then~\cite{McKilliam2009CoxeterLattices} for fast algorithms for all of the Coxeter lattices that includes $A_n$ and $A_n^*$, $E_8$, etc.  Technically, the algorithm from~\cite{McKilliam2009CoxeterLattices} does include that from~\cite{McKilliam2008}.  However~\cite{McKilliam2008} is simpler and shorter to read first.

\item\textbf{Comment:}  
When saying "if the obtuse superbasis is known", you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
\\\textbf{Response:}

### 2014-07-08 paper.tex
minor

The closest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms all require a number of operations that grows as $O(n^{O(n)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such as the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, the integer lattice $\ints^{n}$, and the Leech lattice~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec,Conway1986SoftDecLeechGolay, Clarkson1999:Anstar, McKilliam2008, McKilliam2008b, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.  In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.  Each lattice of Voronoi's first kind has what is called an \emph{obtuse superbasis}.  We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.  This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.  Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.  We show that the series converges to a closest lattice point after at most $n$ terms, resulting in $O(n^4)$ operations in total.  This result extends upon a recent result by some of the authors showing that a short vector in a lattice of Voronoi's first kind can be found by computing a minimum cut in a weighted graph~\cite{McKilliam_short_vectors_first_type_isit_2012}.

>>>>

The closest lattice point problem is known to be NP-hard in general~\cite{micciancio_hardness_2001, Dinur2003_approximating_CVP_NP_hard, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a closest lattice point in reasonable time if the dimension is small~\cite{Pohst_sphere_decoder_1981,Kannan1987_fast_general_np,Agrell2002}.  These algorithms require a number of operations that grows as $O(n^{O(n)})$ or $O(n^{O(n^2)})$ where $n$ is the dimension of the lattice.  Recently, Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} described a solution for the closest lattice point problem that requires a number of operations that grows as $O(2^{2n})$.  This single exponential growth in complexity is the best known.

Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.  In this paper we consider a particular class of lattices, those of \emph{Voronoi's first kind}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases,Valentin2003_coverings_tilings_low_dimension,Voronoi1908_main_paper}.  Each lattice of Voronoi's first kind has what is called an \emph{obtuse superbasis}.  We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.  This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.  Each relevant vector in the series can be computed in $O(n^3)$ operations using an algorithm for computing a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.  We show that the series converges to a closest lattice point after at most $n$ terms, resulting in $O(n^4)$ operations in total.  This result extends upon a recent result by some of the authors showing that a short vector in a lattice of Voronoi's first kind can be found by computing a minimum cut in a weighted graph~\cite{McKilliam_short_vectors_first_type_isit_2012}.

%BLERG incase I do want to introduce graph Laplacian in the introduction.
%We show that if the obtuse superbasis is known, then a closest lattice point can be computed in $O(n^4)$ operations.  This is achieved by enumerating a series of \emph{relevant vectors} of the lattice.  We show that the series converges to a closest lattice point after at most $n$ terms and, using a connection between the obtuse superbasis and the Laplacian of a graph with positive weights, we show how each relevant vector in the series can be computed in $O(n^3)$ operations by an algorithm that computes a minimum cut in an undirected flow network~\cite{Picard_min_cuts_1974,Sankaran_solving_CDMA_mincut_1998,Ulukus_cdma_mincut_1998,Cormen2001}.
